### 2020-06-03 report/sections/server/node/efsm.tex
Report: progress on chapter 4

@@

\begin{itemize}
\item callbacks -- type aliases for each state (handler)
\item need a way for the runtime to distinguish between states (implementation)
\item attempt at using conditional types, but because they handle conditional union types in a distributed way (give example), cannot be statically typed
\item alternative -- use discriminated union by wrapping in an Implementation class
\item all non-terminal states will `return' the successor implementation -- very difficult to resolve in the type-checker in the runtime (give example of what it may look like) -- solve by giving each state implementation the `advance' function to respect the event-driven nature of everything, so they can call it on completion
\item visualise the interaction between runtime and states as a sequence diagram for message passing
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
// Inside the Message namespace...
export type S54THANKS = {
	label: Labels.S54.THANKS,
	payload: [string],
};
export type S54TERMINATE = {
	label: Labels.S54.TERMINATE,
	payload: [],
};

export type S54 = | S54THANKS | S54TERMINATE;
\end{lstlisting}
%\label{}
\captionof{lstlisting}{Generated Message Type Definition for State 54}
\end{figure}

@@

We collect the APIs that the developer needs to implement
under the \texttt{Handler} namespace. 
As a design choice, we \textit{do not} generate handlers for
terminal states, because the semantics of inactivity mean
there is nothing to handle.
We outline how the generated API differs between sending
and receiving states.

\paragraph{Send}
We generalise deterministic send actions as a trivial \textit{selection}, 
as motivated from the theory (\dots).

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
// Inside the Handler namespace...
export type S54 = 
	| [Labels.S54.THANKS, Message.S54THANKS['payload'],
		Implementation.S52] 
	| [Labels.S54.TERMINATE, Message.S54TERMINATE['payload'], 
		Implementation.S52];
\end{lstlisting}
\captionof{lstlisting}{Generated Type for \trole{Svr} Send State
in \tprotocol{Adder} protocol}
\end{figure}

\paragraph{Receive}
\begin{itemize}
\item very short -- go over the structure and give examples
\item show how it extends for polyadic payloads
\end{itemize}

@@

As with send states,
we generalise deterministic receive actions as a trivial \textit{branch}, 
as motivated from the theory (\dots).

\subsection{Terminal}
\begin{itemize}
\item basically nothing here -- can highlight the provided type alias to provide a better developer experience
\end{itemize}

>>>>

We show the structure of the generated EFSM.ts file in
\cref{lst:nodeefsmfile}.
Note that the formal definition of the EFSM in 
\cref{section:scribbleefsm}
contains more than just states and the state transition function,
so we encode the additional information as well.
Each type of information is grouped into their own
\textit{namespace}, and are collectively exported in
the EFSM \textit{module} for the developer to use.

@@

\begin{itemize}
\item callbacks -- type aliases for each state (handler)
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
// Inside the Message namespace...
export interface S54THANKS {
	label: Labels.S54.THANKS,
	payload: [string],
};
export interface S54TERMINATE {
	label: Labels.S54.TERMINATE,
	payload: [],
};

export type S54 = | S54THANKS | S54TERMINATE;
\end{lstlisting}
\label{lst:addersvrmsg}
\captionof{lstlisting}{Generated Message Type Definition for State 54}
\end{figure}

By expressing the payload type as a \textit{tuple}\footnote{
In TypeScript, a tuple is an array with fixed size
and known types for elements at each position.
},
we easily generalise our type definition to polyadic payloads.

@@

We collect the APIs that the developer needs to implement
under the \texttt{Handler} namespace. 
As a design choice, we \textit{do not} generate handlers for
terminal states, because the semantics of inactivity mean
there is nothing to handle.
%Hence, all generated handler APIs are for non-terminal states,
%and will hold some reference to the successor state encoding.
%The reader will notice that, in the listings below, the successor
%state is stated to be under the 
We outline how the generated API differs between sending
and receiving states.

\subparagraph{Send}
We model selections using \textit{union types} to
encapsulate the possible send actions. 
Each send action is encoded as a tuple of
the label, the payload, and the successor state.
We see some benefits from defining Message Types as interfaces:
with deterministic properties, we can index into the message type
to access the property's type, 
e.g. \texttt{Message.S54THANKS['payload']} would resolve to
\texttt{[string]}, based on \cref{lst:addersvrmsg}.
We provide an example in \cref{lst:addersvrsendhandler}.

We generalise deterministic send actions as a trivial \textit{selection}, 
as motivated from the theory (\cref{fig:globaltypes}),
so the encoding for State 53 in the \trole{Svr} FSM would be
the union of a single tuple.

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
// Inside the Handler namespace...
export type S54 = 
	| [Labels.S54.THANKS, Message.S54THANKS['payload'],
		Implementation.S52] 
	| [Labels.S54.TERMINATE, Message.S54TERMINATE['payload'], 
		Implementation.S52];
\end{lstlisting}
\captionof{lstlisting}{Generated Type for \trole{Svr} Send State
in \tprotocol{Adder} protocol}
\label{lst:addersvrsendhandler}
\end{figure}

\subparagraph{Receive}
\begin{itemize}
\item very short -- go over the structure and give examples
\item show how it extends for polyadic payloads
\end{itemize}

@@

As with send states,
we generalise deterministic receive actions as a trivial \textit{branch},
which would be an interface with one property.

\subsection{Wrapping Handlers in ``Implementations''}
\begin{itemize}
\item need a way for the runtime to distinguish between states (implementation)
\item attempt at using conditional types, but because they handle conditional union types in a distributed way (give example), cannot be statically typed
\item alternative -- use discriminated union by wrapping in an Implementation class
\item all non-terminal states will `return' the successor implementation -- very difficult to resolve in the type-checker in the runtime (give example of what it may look like) -- solve by giving each state implementation the `advance' function to respect the event-driven nature of everything, so they can call it on completion
\item visualise the interaction between runtime and states as a sequence diagram for message passing
\end{itemize}
